### 2024-10-08 paper.tex
First commit

\documentclass[11pt,sigconf]{iabart}
%\documentclass[10pt,sigconf,letterpaper]{acmart}
%\usepackage[vskip=1em,font=itshape,leftmargin=2em,rightmargin=2em]{quoting}
\usepackage[skip=4pt plus1pt]{parskip}
%\copyrightyear{2024}
%\acmYear{2024}
%\setcopyright{acmlicensed}\acmConference[IAB Network Management WS]{The IAB Next Era of Network Management Operations}{December 3, 2024}{}
%\acmBooktitle{The IAB next Era of Network Management Workshop}{December 3, 2024}{}

\begin{document}

\title{LwM2M TITLE WIP}

\author{Jaime Jiménez}
\email{jaime.jimenez@ericsson.com}
\orcid{0009-0008-2864-4269}
\affiliation{%
  \institution{Ericsson}
  \city{Jorvas}
  \country{Finland}
}
\author{Matthew Gillmore}
\email{email@email}
\orcid{orcidID}
\affiliation{%
  \institution{Itron}
  \city{City}
  \country{Country}
}

\begin{abstract}
%Jaime: Write a concise abstract (150-250 words) summarizing the paper's purpose, methods, key findings, and implications.
TBD.

\end{abstract}

\keywords{IoT, network management, LwM2M, device management, protocol, scalability, security}

\maketitle

\section{Introduction} \label{introduction}

%Jaime: We need to address gaps, such as operator-specific challenges or more detailed recommendations for future standardization efforts.

The rapid evolution of network management protocols necessitates a reevaluation of existing technologies and their applicability to modern challenges. The Lightweight Machine-to-Machine (LwM2M) protocol~\cite{lwm2m-spec}, developed by the Open Mobile Alliance (OMA), is a key player in this domain, offering a standardized framework for managing Internet of Things (IoT) devices~\cite{oma-sdo}. This paper explores the role of LwM2M in the context of the IAB workshop on the Next Era of Network Management Operations, focusing on its current deployments, challenges, and future potential.

The IAB "NEMOPS" workshop seeks contributions that critically assess the progress made since the 2002 IAB workshop, particularly in terms of network management protocols. This paper aims to present LwM2M's as a new protocol created after 2002 that addresses the needs for managing IoT endpoints from the operational point of view of device and network management.

Our contribution is informed by the authors' extensive experience with IoT and contributions in the IoT domain both at IETF and in OMA. The rest of the document is organized as follows: The \hyperref[introduction]{Introduction} outlines LwM2M as a standardized framework for managing IoT devices, addressing current network management challenges. The \hyperref[overview]{LwM2M Protocol Overview} details its architecture, focusing on communication between Clients, Servers, and Bootstrap Servers (see Figure~\ref{fig:overall_architecture}). Recent advancements, including integrations with blockchain and industrial protocols, are discussed in \hyperref[extensions]{LwM2M Extensions}, highlighting its adaptability. Remaining adoption challenges are explored in \hyperref[operations]{LwM2M Operations}, while the \hyperref[conclusions]{Conclusions} summarize key findings and future directions.

\textbf{LwM2M and IETF}

The IETF has played a fundamental role in shaping the protocols that underpin Lightweight Machine-to-Machine (LwM2M). IETF efforts have fcused on adapting existing Internet and Web protocols to meet the needs of resource-constrained IoT devices ~\cite{9139045}. These protocols enable efficient communication, security, and interoperability, forming a foundation for LwM2M.

The Lightweight Machine-to-Machine (LwM2M) protocol is built upon several key IETF standards. At the core is the \textit{Constrained Application Protocol (CoAP)}~\cite{rfc7252}, a lightweight RESTful protocol designed for constrained environments, providing the fundamental request/ response model for LwM2M communications. CoAP supports features such as observe/notify, which are critical for resource updates. Additionally, \textit{RFC 7959}~\cite{rfc7959} defines block-wise transfers in CoAP, allowing LwM2M to efficiently handle large payloads by breaking them into smaller blocks. The \textit{LwM2M} protocol also leverages \textit{RFC 7641}~\cite{rfc7641} for resource observation, enabling clients to monitor changes to resources without continuous polling. For secure communications, LwM2M often relies on the \textit{Datagram Transport Layer Security (DTLS)} as outlined in \textit{RFC 6347}~\cite{rfc6347}, ensuring encryption and integrity over the CoAP protocol. An additional layer of security is provided by \textit{Object Security for Constrained RESTful Environments (OSCORE)}~\cite{rfc8613}, which offers end-to-end encryption and integrity protection directly at the application layer, making it suitable for scenarios where DTLS is not applicable.

Furthermore, LwM2M also supports other transport protocols such as \textit{MQTT}~\cite{mqtt-spec} and \textit{HTTP}~\cite{http-spec}, broadening its applicability across different network environments and use cases.

\section{LwM2M Protocol Overview} \label{overview}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{figs/arch.pdf}
  \caption{General LwM2M Architecture}
  \label{fig:overall_architecture}
\end{figure}

This section presents an overview of the LwM2M protocol, emphasizing its participating entities, data model, and communication interfaces. Additionally, it describes the LwM2M library's design and functionality, highlighting challenges encountered during its development.

\textbf{Participating Entities}

The LwM2M protocol defines three primary entities that form the backbone of its communication architecture:
\begin{itemize}
    \item \textbf{LwM2M Client}: This entity is typically a device responsible for data collection and managing resources, such as sensors or other IoT devices. The client manages the lifecycle of various objects and interacts with servers to share collected data and receive management commands.
    \item \textbf{LwM2M Server}: This server is responsible for managing multiple clients, aggregating their data, and issuing commands for configuration or maintenance. It plays a crucial role in the control and coordination of client devices within an IoT ecosystem.
    \item \textbf{LwM2M Bootstrap Server}: This unique entity is tasked with the initial configuration of the LwM2M Client. Unlike other protocols, LwM2M introduces the Bootstrap Server to facilitate the initial setup of devices, especially when multiple servers are involved. During the bootstrap process, the Bootstrap Server provides the client with configuration details, such as security credentials and connection information, which can either be pre-integrated into the device's software or dynamically provided by the Bootstrap Server \cite{pop00006}.
\end{itemize}

The introduction of the Bootstrap Server differentiates LwM2M from other IoT protocols. Before the client can establish a connection to a server, it undergoes the bootstrap procedure to load initial configurations. This feature is particularly advantageous in scenarios with multiple servers or when load balancing is required, as it allows for flexible and dynamic configuration \cite{pop00006}.

\textbf{LwM2M Data Model}

The LwM2M protocol utilizes a structured, object-based data model, which simplifies communication between clients and servers. Each data entity in the LwM2M model is defined as an object, identified by a unique integer ID, as specified by the Open Mobile Alliance (OMA). These objects can represent various resources such as sensors, actuators, or configuration settings.

Each object comprises multiple resources, which are the fundamental data points within the object. Resources are also assigned integer IDs and are categorized as mandatory or optional based on the object's intended use. In some cases, an object can have multiple instances, each with a unique instance ID. This flexibility allows for scenarios where, for example, a device might have several temperature sensors, each represented as an instance of the same object \cite{pop00006}.

The LwM2M data model enables servers to access individual resources, instances, or entire objects using well-defined URI strings. The format of these URIs is as follows:
\[
/<Object ID>/<Instance ID>/<Resource ID>
\]
In this structure, the instance or resource ID can be omitted if the request targets the entire object or a specific instance. This model provides a straightforward way to interact with data at different levels of granularity.

\textbf{Communication Interfaces}

LwM2M defines four primary communication interfaces that facilitate interactions between clients and servers, each serving distinct roles in the protocol's operation:

\begin{itemize}
    \item \textbf{Bootstrap Interface}: This interface manages the bootstrap procedure, allowing the client to acquire initial configurations and security credentials from the Bootstrap Server before connecting to an LwM2M Server. The process can be automated or initiated by the client as needed.
    \item \textbf{Registration Interface}: After the bootstrap procedure, the client registers with the LwM2M Server through the Registration Interface. During registration, the client provides its endpoint name, which acts as an access token. The server can deny access if the provided token is invalid. To maintain its registration status, the client must periodically send updates. If these updates are not received within the agreed time, the client is considered de-registered and must re-initiate the registration process \cite{pop00006}.
    \item \textbf{Device Management and Service Enablement Interface}: Once registered, the server can use this interface to perform various management tasks, such as reading and writing data or executing commands on the client. It allows the server to control the client's resources, adjusting settings or triggering actions as required.
    \item \textbf{Information Reporting Interface}: This interface enables the client to report changes in its status or resource values to the server, using the Observe function. The server can set up observation requests, prompting the client to notify it when certain resource values change or when predefined conditions are met. This capability is particularly useful for monitoring dynamic IoT environments where timely updates are crucial. LwM2M defines specific attributes to configure how frequently these notifications occur \cite{pop00006}.
\end{itemize}

LwM2M has been integrated on most IoT OSs like Mbed~\cite{mbed}, RIOT OS~\cite{riot_os}, Contiki-NG~\cite{contiki_ng}, and others, with a strong focus on IoT device security~\cite{pop00010}.

\section{LwM2M Extensions} \label{extensions}

LwM2M, while robust in its core functionalities, has seen significant advancements through various extensions and integrations. This section delves into the recent developments and integrations of LwM2M, highlighting its evolving role in the IoT ecosystem.

\subsection{LwM2M Integrations}

%Jaime: Expand on the implications of these integrations for the future of IoT and network management.
The integration of LwM2M with other technologies has been researched, aiming to enhance the capabilities and interoperability of IoT systems.

To name a few, Kim et al.~\cite{pop00007} designed and implemented a blockchain-based system that enables interworking between oneM2M and LwM2M IoT systems. This approach not only improves interoperability between different IoT frameworks but also enhances security by leveraging blockchain's immutable and distributed ledger properties.

In the context of Industrial IoT (IIoT), Yaker et al.~\cite{pop00004} introduced a novel edge Security Information and Event Management (SIEM) system for managing IoT flows within 5G private networks. Their approach incorporates LwM2M data events, demonstrating how edge computing can effectively manage and secure IoT data in a 5G environment. Similarly, Myoung et al.~\cite{pop00012} addressed the integration of LwM2M with smart metering technologies, proposing a data interworking model between the Device Language Message Specification (DLMS)~\cite{dlms-spec} nd LwM2M protocols. Their work highlights the potential for harmonizing different IoT protocols within Advanced Metering Infrastructure (AMI) systems, thereby enabling more efficient data exchange and management.

Further efforts have been made to integrate LwM2M with industrial communication protocols. Karaagac et al.~\cite{pop00013} explored the interoperability between LwM2M and OPC UA (Open Platform Communications Unified Architecture), proposing a framework where OPC UA Servers can be virtualized as LwM2M Clients and vice versa. This approach aims to bridge the gap between IoT and industrial automation, facilitating seamless communication across heterogeneous systems. Similarly, Cavalcanti et al.~\cite{pop00014} reviewed various machine-to-machine communication protocols within the context of Industry 4.0, emphasizing the advantages of integrating OPC UA with LwM2M for industrial applications.

A last example, Wang et al.~\cite{pop00015} proposed a CoAP-based OPC UA transmission scheme tailored for resource-constrained devices. Although their work does not directly involve LwM2M, it illustrates the broader trend of adapting industrial communication protocols to fit the needs of IoT environments. Such efforts are indicative of a growing interest in enabling resource-efficient and interoperable solutions in the IoT landscape, especially where industrial and IoT domains converge.

The integration of LwM2M with these diverse technologies underscores its versatility and adaptability, positioning it as a critical component in the evolving IoT ecosystem.

\subsection{ Adoption of LwM2M in IoT Security}

The Lightweight Machine-to-Machine (LwM2M) protocol has emerged as a critical focus in recent IoT research, particularly concerning security improvements, implementation strategies, and integration with other technologies. This increased attention highlights its growing importance in advancing IoT systems.

Several studies have concentrated on enhancing IoT security through the use of LwM2M. For instance, Muhammad et al.~\cite{pop00001} investigated the use of ARIA cryptography within Hardware Secure Modules for both LwM2M and MQTT protocols, aiming to bolster the security of next-generation IoT systems. Similarly, Lanzieri et al.~\cite{pop00011} proposed extensions to the LwM2M core specification to facilitate secure and authorized client-to-client communication, thereby addressing a significant limitation of the standard protocol.

However, despite its promising capabilities, LwM2M still faces challenges in widespread adoption. Szydlo and Nagy~\cite{pop00003} pointed out the low market uptake of LwM2M in IoT products, attributing this to a lack of freely available developer tools. This observation underscores the need for more accessible resources to facilitate LwM2M implementation.

In parallel, Saif and Matrawy~\cite{pop00005} proposed and evaluated methods for the mass configuration of heterogeneous IIoT nodes using LwM2M, demonstrating potential solutions for scaling the protocol in diverse environments.

\section{LwM2M Operations} \label{operations}
%Jaime: Develop this section to cover typical operations, challenges, and best practices. Include examples of common operations and how they're implemented.
%Jaime: Add more specific examples of how LwM2M is used in real-world applications.

TBD

\section{Conclusions} \label{conclusions}

TBD

\section{Acknowledgments}

We'd like to thank Ericsson for their support of this work. We also appreciate A, B, C for their valuable discussions on this topic.

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}
\endinput

### 2024-10-09 paper.tex
first editorial pass

@@
 The IAB "NEMOPS" workshop seeks contributions that critically assess the progress made since the 2002 IAB workshop, particularly in terms of network management protocols.
-This paper aims to present LwM2M's as a new protocol created after 2002 that addresses the needs for managing IoT endpoints from the operational point of view of device and network management.
+This paper aims to present LwM2M's as a new protocol created in 2017 that addresses the needs for managing IoT endpoints from the operational point of view of device and network management.
 
@@
 The IETF has played a fundamental role in shaping the protocols that underpin Lightweight Machine-to-Machine (LwM2M).
-IETF efforts have fcused on adapting existing Internet and Web protocols to meet the needs of resource-constrained IoT devices ~\cite{9139045}.
-These protocols enable efficient communication, security, and interoperability, forming a foundation for LwM2M.
+IETF efforts have focused on adapting existing Internet and Web protocols to meet the needs of resource-constrained IoT devices ~\cite{9139045}.
 
-The Lightweight Machine-to-Machine (LwM2M) protocol is built upon several key IETF standards.
-At the core is the \textit{Constrained Application Protocol (CoAP)}~\cite{rfc7252}, a lightweight RESTful protocol designed for constrained environments, providing the fundamental request/ response model for LwM2M communications.
-CoAP supports features such as observe/notify, which are critical for resource updates.
-Additionally, \textit{RFC 7959}~\cite{rfc7959} defines block-wise transfers in CoAP, allowing LwM2M to efficiently handle large payloads by breaking them into smaller blocks.
-The \textit{LwM2M} protocol also leverages \textit{RFC 7641}~\cite{rfc7641} for resource observation, enabling clients to monitor changes to resources without continuous polling.
+LwM2M is built upon several key IETF standards.
+At the core is the \textit{Constrained Application Protocol (CoAP)}~\cite{rfc7252}, a lightweight RESTful protocol designed for constrained environments, providing the fundamental request/response model for LwM2M communications.
+\textit{RFC 7959}~\cite{rfc7959} defines block-wise transfers in CoAP, allowing LwM2M to efficiently handle large payloads by breaking them into smaller blocks.
+The \textit{LwM2M} protocol also leverages \textit{RFC 7641}~\cite{rfc7641} for resource observation, enabling clients to subscribe to changes of resource state without continuous polling.
 For secure communications, LwM2M often relies on the \textit{Datagram Transport Layer Security (DTLS)} as outlined in \textit{RFC 6347}~\cite{rfc6347}, ensuring encryption and integrity over the CoAP protocol.
 An additional layer of security is provided by \textit{Object Security for Constrained RESTful Environments (OSCORE)}~\cite{rfc8613}, which offers end-to-end encryption and integrity protection directly at the application layer, making it suitable for scenarios where DTLS is not applicable.
+Additionally, the \textit{Constrained RESTful Environments (CoRE) Resource Directory (RD)}~\cite{rfc9176} facilitates resource registration of IoT endpoints by maintaining information about resources on other servers, the lookup interface, although present is not intended for applications but for management purposes.
+\sloppy
 
-Furthermore, LwM2M also supports other transport protocols such as \textit{MQTT}~\cite{mqtt-spec} and \textit{HTTP}~\cite{http-spec}, broadening its applicability across different network environments and use cases.
+Furthermore, LwM2M supports additional transport protocols, including the \textit{Hypertext Transfer Protocol (HTTP)}~\cite{http-spec} and \textit{Message Queuing Telemetry Transport (MQTT)}~\cite{mqtt-spec}, the latter developed by the OASIS foundation~\cite{oasis-foundation}.
+This broadens its applicability across various network environments and use cases.
 
@@
 The LwM2M protocol defines three primary entities that form the backbone of its communication architecture:
+
 \begin{itemize}
-    \item \textbf{LwM2M Client}: This entity is typically a device responsible for data collection and managing resources, such as sensors or other IoT devices.
-The client manages the lifecycle of various objects and interacts with servers to share collected data and receive management commands.
-    \item \textbf{LwM2M Server}: This server is responsible for managing multiple clients, aggregating their data, and issuing commands for configuration or maintenance.
-It plays a crucial role in the control and coordination of client devices within an IoT ecosystem.
-    \item \textbf{LwM2M Bootstrap Server}: This unique entity is tasked with the initial configuration of the LwM2M Client.
-Unlike other protocols, LwM2M introduces the Bootstrap Server to facilitate the initial setup of devices, especially when multiple servers are involved.
-During the bootstrap process, the Bootstrap Server provides the client with configuration details, such as security credentials and connection information, which can either be pre-integrated into the device's software or dynamically provided by the Bootstrap Server \cite{pop00006}.
+  \item \textbf{LwM2M Client}: Typically a smart device, such as a light, smart meter, switch, or gateway, responsible for data collection and resource management at the device level.
+The client oversees the lifecycle of various objects and communicates with servers to transmit data and receive management instructions.
+  \item \textbf{LwM2M Server}: Often referred to as the manager, this server manages multiple clients, aggregates their data, and issues configuration or maintenance commands.
+It is essential for controlling and coordinating client devices within an IoT ecosystem.
+  \item \textbf{LwM2M Bootstrap Server}: This entity is responsible for the initial configuration of the LwM2M Client.
+Unlike other protocols, LwM2M includes a Bootstrap Server to streamline device setup, particularly when multiple servers are involved.
+During the bootstrap process, it provides the client with configuration details, such as security credentials and connection information, which can be pre-integrated into the device's software or dynamically supplied by the Bootstrap Server.
 \end{itemize}
@@
 The introduction of the Bootstrap Server differentiates LwM2M from other IoT protocols.
-Before the client can establish a connection to a server, it undergoes the bootstrap procedure to load initial configurations.
-This feature is particularly advantageous in scenarios with multiple servers or when load balancing is required, as it allows for flexible and dynamic configuration \cite{pop00006}.
+Before the client can establish a connection to a server, it undergoes the bootstrap procedure to load initial configurations and key material.
+This feature is particularly advantageous in scenarios with multiple servers or when load balancing is required, as it allows for flexible and dynamic configuration.
 
@@
 
-The LwM2M protocol utilizes a structured, object-based data model, which simplifies communication between clients and servers.
-Each data entity in the LwM2M model is defined as an object, identified by a unique integer ID, as specified by the Open Mobile Alliance (OMA).
-These objects can represent various resources such as sensors, actuators, or configuration settings.
+The LwM2M protocol employs a structured, object-based data model to facilitate communication between clients and servers.
+In this model, each data entity is defined as an object, uniquely identified by an integer ID, as specified by the Open Mobile Alliance (OMA).
+These objects can represent various resources, such as sensors, actuators, or configuration settings.
 
-Each object comprises multiple resources, which are the fundamental data points within the object.
-Resources are also assigned integer IDs and are categorized as mandatory or optional based on the object's intended use.
-In some cases, an object can have multiple instances, each with a unique instance ID.
-This flexibility allows for scenarios where, for example, a device might have several temperature sensors, each represented as an instance of the same object \cite{pop00006}.
+Each object consists of multiple resources, which serve as the fundamental data points within the object.
+Resources are assigned integer IDs and are classified as either mandatory or optional, depending on the object's intended function being the mandatory ones used for interoperability purposes.
+An object can also have multiple instances, each with a unique instance ID, allowing for flexibility.
+For instance, a device with several sensors can represent each sensor as an instance of the same object, such as in a network of connected lightbulbs.
 
 The LwM2M data model enables servers to access individual resources, instances, or entire objects using well-defined URI strings. The format of these URIs is as follows:
+
 \[
@@
 \]
+
 In this structure, the instance or resource ID can be omitted if the request targets the entire object or a specific instance.
 This model provides a straightforward way to interact with data at different levels of granularity.
+
+All objects are listed in the LwM2M Registry~\cite{lwm2m-registry}, categorized as either application-oriented (e.g., thermostats, lamps) or system-oriented (e.g., security, connectivity).
 
@@
 \begin{itemize}
-    \item \textbf{Bootstrap Interface}: This interface manages the bootstrap procedure, allowing the client to acquire initial configurations and security credentials from the Bootstrap Server before connecting to an LwM2M Server.
-The process can be automated or initiated by the client as needed.
+  \item \textbf{Bootstrap Interface}: This interface facilitates the initial setup by enabling the client to obtain necessary configurations and security credentials from the Bootstrap Server prior to connecting with an LwM2M Server.
+The process can be either automated or client-initiated.
+
   \item \textbf{Registration Interface}: After the bootstrap procedure, the client registers with the LwM2M Server through the Registration Interface.
@@
 To maintain its registration status, the client must periodically send updates.
-If these updates are not received within the agreed time, the client is considered de-registered and must re-initiate the registration process \cite{pop00006}.
+If these updates are not received within the agreed time, the client is considered de-registered and must re-initiate the registration process.
+
   \item \textbf{Device Management and Service Enablement Interface}: Once registered, the server can use this interface to perform various management tasks, such as reading and writing data or executing commands on the client.
 It allows the server to control the client's resources, adjusting settings or triggering actions as required.
+
   \item \textbf{Information Reporting Interface}: This interface enables the client to report changes in its status or resource values to the server, using the Observe function.
@@
 This capability is particularly useful for monitoring dynamic IoT environments where timely updates are crucial.
-LwM2M defines specific attributes to configure how frequently these notifications occur \cite{pop00006}.
+LwM2M defines specific attributes to configure how frequently these notifications occur.
 \end{itemize}
@@
 
-LwM2M, while robust in its core functionalities, has seen significant advancements through various extensions and integrations.
-This section delves into the recent developments and integrations of LwM2M, highlighting its evolving role in the IoT ecosystem.
+LwM2M, though robust in its core functionalities, has advanced significantly through various extensions and integrations.
+This section explores recent research developments that may inspire its future evolution.
 

### 2024-10-09 paper.tex
second editorial pass

@@
 
-The IETF has played a fundamental role in shaping the protocols that underpin Lightweight Machine-to-Machine (LwM2M).
+The IETF has played a fundamental role in shaping the protocols that underpin LwM2M.
 IETF efforts have focused on adapting existing Internet and Web protocols to meet the needs of resource-constrained IoT devices ~\cite{9139045}.
@@
 
-LwM2M has been integrated on most IoT OSs like Mbed~\cite{mbed}, RIOT OS~\cite{riot_os}, Contiki-NG~\cite{contiki_ng}, and others, with a strong focus on IoT device security~\cite{pop00010}.
+LwM2M has been integrated on most IoT OSs like Mbed~\cite{mbed}, RIOT OS~\cite{riot_os}, Contiki-NG~\cite{contiki_ng}, and many others, with a strong focus on IoT device security~\cite{pop00010}.
 
-\section{LwM2M Extensions} \label{extensions}
+\section{LwM2M Evolution} \label{extensions}
 
@@
 This section explores recent research developments that may inspire its future evolution.
+%Jaime: Expand on the implications of these integrations for the future of IoT and network management.
 
-\subsection{LwM2M Integrations}
+LwM2M has become a central focus in IoT research due to its improved security, in particular for firmware updates ~\cite{8725488} and integration with other technologies.
 
-%Jaime: Expand on the implications of these integrations for the future of IoT and network management.
-The integration of LwM2M with other technologies has been researched, aiming to enhance the capabilities and interoperability of IoT systems.
+Several studies have concentrated on enhancing IoT security through the use of LwM2M.
+For instance, Muhammad et al.~\cite{pop00001} investigated the use of ARIA cryptography within Hardware Secure Modules for both LwM2M and MQTT protocols, aiming to bolster the security of next-generation IoT systems.
+Similarly, Lanzieri et al.~\cite{pop00011} proposed extensions to the LwM2M core specification to facilitate secure and authorized client-to-client communication, thereby addressing a significant limitation of the current standard.
 
-To name a few, Kim et al.~\cite{pop00007} designed and implemented a blockchain-based system that enables interworking between oneM2M and LwM2M IoT systems.
+Others have focused on interoperability between SDOs, Kim et al.~\cite{pop00007} have designed and implemented a blockchain-based system that enables interworking between oneM2M~\cite{oneM2M-spec} and LwM2M IoT systems.
 This approach not only improves interoperability between different IoT frameworks but also enhances security by leveraging blockchain's immutable and distributed ledger properties.
@@
 Their approach incorporates LwM2M data events, demonstrating how edge computing can effectively manage and secure IoT data in a 5G environment.
-Similarly, Myoung et al.~\cite{pop00012} addressed the integration of LwM2M with smart metering technologies, proposing a data interworking model between the Device Language Message Specification (DLMS)~\cite{dlms-spec} nd LwM2M protocols.
-Their work highlights the potential for harmonizing different IoT protocols within Advanced Metering Infrastructure (AMI) systems, thereby enabling more efficient data exchange and management.
+Similarly, Myoung et al.~\cite{pop00012} addressed the integration of LwM2M with smart metering technologies, proposing a data interworking model between the Device Language Message Specification (DLMS)~\cite{dlms-spec} used for managing smart meters and the LwM2M protocol.
+This highlights the potential for harmonizing different IoT protocols within metering infrastructure systems.
 
 Further efforts have been made to integrate LwM2M with industrial communication protocols.
-Karaagac et al.~\cite{pop00013} explored the interoperability between LwM2M and OPC UA (Open Platform Communications Unified Architecture), proposing a framework where OPC UA Servers can be virtualized as LwM2M Clients and vice versa.
+Karaagac et al.~\cite{pop00013} explored the interoperability between LwM2M and the Open Platform Communications Unified Architecture (OPC UA), proposing a framework where OPC UA Servers can be virtualized as LwM2M Clients and vice versa.
 This approach aims to bridge the gap between IoT and industrial automation, facilitating seamless communication across heterogeneous systems.
@@
 A last example, Wang et al.~\cite{pop00015} proposed a CoAP-based OPC UA transmission scheme tailored for resource-constrained devices.
-Although their work does not directly involve LwM2M, it illustrates the broader trend of adapting industrial communication protocols to fit the needs of IoT environments.
-Such efforts are indicative of a growing interest in enabling resource-efficient and interoperable solutions in the IoT landscape, especially where industrial and IoT domains converge.
-
-The integration of LwM2M with these diverse technologies underscores its versatility and adaptability, positioning it as a critical component in the evolving IoT ecosystem.
-
-\subsection{ Adoption of LwM2M in IoT Security}
-
-The Lightweight Machine-to-Machine (LwM2M) protocol has emerged as a critical focus in recent IoT research, particularly concerning security improvements, implementation strategies, and integration with other technologies.
-This increased attention highlights its growing importance in advancing IoT systems.
-
-Several studies have concentrated on enhancing IoT security through the use of LwM2M.
-For instance, Muhammad et al.~\cite{pop00001} investigated the use of ARIA cryptography within Hardware Secure Modules for both LwM2M and MQTT protocols, aiming to bolster the security of next-generation IoT systems.
-Similarly, Lanzieri et al.~\cite{pop00011} proposed extensions to the LwM2M core specification to facilitate secure and authorized client-to-client communication, thereby addressing a significant limitation of the standard protocol.
-
-However, despite its promising capabilities, LwM2M still faces challenges in widespread adoption.
-Szydlo and Nagy~\cite{pop00003} pointed out the low market uptake of LwM2M in IoT products, attributing this to a lack of freely available developer tools.
-This observation underscores the need for more accessible resources to facilitate LwM2M implementation.
-
-In parallel, Saif and Matrawy~\cite{pop00005} proposed and evaluated methods for the mass configuration of heterogeneous IIoT nodes using LwM2M, demonstrating potential solutions for scaling the protocol in diverse environments.
+Although their work does not directly involve LwM2M, it illustrates the broader trend of adapting industrial communication protocols to fit the needs of IoT environments and the reuse of IETF protocols in wider contexts that originally intended.
+This could also indicate the interest in enabling resource-efficient and interoperable solutions in the IoT landscape, where industrial and IoT domains converge.
 
@@
 
-We'd like to thank Ericsson for their support of this work. We also appreciate A, B, C for their valuable discussions on this topic.
+We'd like to thank XYZ for their support of this work. We also appreciate A, B, C for their valuable discussions on this topic.
 

### 2024-10-16 paper.tex
pre-review version

@@
 }
-\author{Matthew Gillmore}
-\email{email@email}
-\orcid{orcidID}
-\affiliation{%
-  \institution{Itron}
-  \city{City}
-  \country{Country}
-}
 
 \begin{abstract}
-%Jaime: Write a concise abstract (150-250 words) summarizing the paper's purpose, methods, key findings, and implications.
-TBD.
+
+This paper provides an overview of the Lightweight Machine-to-Machine (LwM2M) protocol, highlighting its role in IoT device management and its integration with key IETF standards such as CoAP, DTLS, and OSCORE.
+While network management protocols like SNMP, RESTCONF, and CORECONF focus on monitoring and configuring endpoints within the network, LwM2M extends these capabilities by managing the entire lifecycle of individual devices, including configuration, control, and maintenance.
+We discuss LwM2M's architecture, data model, communication interfaces, and explore recent advancements in the protocol.
 
@@
 
-\keywords{IoT, network management, LwM2M, device management, protocol, scalability, security}
+\keywords{Internet of Things, IoT, network management, device management, security, standards}
 
@@
 The IAB "NEMOPS" workshop seeks contributions that critically assess the progress made since the 2002 IAB workshop, particularly in terms of network management protocols.
-This paper aims to present LwM2M's as a new protocol created in 2017 that addresses the needs for managing IoT endpoints from the operational point of view of device and network management.
+This paper aims to present LwM2M, an management protocol that addresses the needs for managing IoT endpoints from the operational point of view of device and network management.
 
@@
 
-The LwM2M data model enables servers to access individual resources, instances, or entire objects using well-defined URI strings. The format of these URIs is as follows:
+The LwM2M data model enables servers to access individual resources, instances, or entire objects using well-defined URI strings.
+The URI template approach follows the Web Linking and the IETF CoRE Link Format~\cite{rfc6690}.
+The format is as follows:
 
@@
 All objects are listed in the LwM2M Registry~\cite{lwm2m-registry}, categorized as either application-oriented (e.g., thermostats, lamps) or system-oriented (e.g., security, connectivity).
+Those for applications follow the same schema and come from the IP for Smart Objects (IPSO)~\cite{ipso} consortium which was integrated into OMA.
+
+Semantically, the object type represents a single measurement, actuation, or control point for example a temperature sensor, a light (actuator), or an on-off switch (control point).
+A resource specifies a particular view or active property of an object.
+For example, a temperature sensor object might expose the current value (most recent reading), also the minimum and maximum possible reading, the minimum and maximum reading in an interval, and attributes like engineering units and application type.
+Attributes describe the metadata configuration, settings, and state of an object or resource, and are discoverable by reading the link-format data of an object or resource.
+Multiple attributes may be serialized in the link-format descriptors that an object exposes.
 
@@
 This section explores recent research developments that may inspire its future evolution.
-%Jaime: Expand on the implications of these integrations for the future of IoT and network management.
-
 LwM2M has become a central focus in IoT research due to its improved security, in particular for firmware updates ~\cite{8725488} and integration with other technologies.
@@
 
-\section{LwM2M Operations} \label{operations}
-%Jaime: Develop this section to cover typical operations, challenges, and best practices. Include examples of common operations and how they're implemented.
-%Jaime: Add more specific examples of how LwM2M is used in real-world applications.
-
-TBD
-
 \section{Conclusions} \label{conclusions}
 
-TBD
+In this paper, we have presented an overview of LwM2M and its components, the dependencies on IETF standards and some potential new building blocks for future additions.
 
@@
 
-We'd like to thank XYZ for their support of this work. We also appreciate A, B, C for their valuable discussions on this topic.
+We'd like to thank Ericsson for their support of this work.
 

### 2024-10-16 paper.tex
camera-ready

@@
 
-\title{LwM2M TITLE WIP}
+\title{Managing IoT Devices with LwM2M}
 
@@
 
-This paper provides an overview of the Lightweight Machine-to-Machine (LwM2M) protocol, highlighting its role in IoT device management and its integration with key IETF standards such as CoAP, DTLS, and OSCORE.
+In this paper we provide an overview of the Lightweight Machine-to-Machine (LwM2M) protocol, a standard device and service management protocol built on top of CoAP for remote management and configuration of devices.
+We highlight its role in IoT device management and its use of key IETF standards such as CoAP, DTLS,and OSCORE.
 While network management protocols like SNMP, RESTCONF, and CORECONF focus on monitoring and configuring endpoints within the network, LwM2M extends these capabilities by managing the entire lifecycle of individual devices, including configuration, control, and maintenance.
-We discuss LwM2M's architecture, data model, communication interfaces, and explore recent advancements in the protocol.
+We also discuss LwM2M's architecture, data model and communication interfaces.
+Last, we also introduce recent advancements in the protocol.
 
@@
 The \hyperref[overview]{LwM2M Protocol Overview} details its architecture, focusing on communication between Clients, Servers, and Bootstrap Servers (see Figure~\ref{fig:overall_architecture}).
-Recent advancements, including integrations with blockchain and industrial protocols, are discussed in \hyperref[extensions]{LwM2M Extensions}, highlighting its adaptability.
-Remaining adoption challenges are explored in \hyperref[operations]{LwM2M Operations}, while the \hyperref[conclusions]{Conclusions} summarize key findings and future directions.
+Recent advancements, including integrations with blockchain and industrial protocols, are discussed in \hyperref[extensions]{LwM2M Extensions}, the \hyperref[conclusions]{Conclusions} summarize the main points.
 
@@
 This section presents an overview of the LwM2M protocol, emphasizing its participating entities, data model, and communication interfaces.
-Additionally, it describes the LwM2M library's design and functionality, highlighting challenges encountered during its development.
 
@@
   \item \textbf{LwM2M Client}: Typically a smart device, such as a light, smart meter, switch, or gateway, responsible for data collection and resource management at the device level.
-The client oversees the lifecycle of various objects and communicates with servers to transmit data and receive management instructions.
+The client oversees the lifecycle of various objects residing in it and communicates with servers to transmit data and receive management instructions.
   \item \textbf{LwM2M Server}: Often referred to as the manager, this server manages multiple clients, aggregates their data, and issues configuration or maintenance commands.
@@
 
-Semantically, the object type represents a single measurement, actuation, or control point for example a temperature sensor, a light (actuator), or an on-off switch (control point).
-A resource specifies a particular view or active property of an object.
-For example, a temperature sensor object might expose the current value (most recent reading), also the minimum and maximum possible reading, the minimum and maximum reading in an interval, and attributes like engineering units and application type.
-Attributes describe the metadata configuration, settings, and state of an object or resource, and are discoverable by reading the link-format data of an object or resource.
-Multiple attributes may be serialized in the link-format descriptors that an object exposes.
-
 \textbf{Communication Interfaces}
@@
   \item \textbf{Registration Interface}: After the bootstrap procedure, the client registers with the LwM2M Server through the Registration Interface.
-During registration, the client provides its endpoint name, which acts as an access token.
-The server can deny access if the provided token is invalid.
+During registration, the client provides its endpoint name whcih together with the security credentials is used as identifier.
 To maintain its registration status, the client must periodically send updates.
@@
 
-LwM2M has been integrated on most IoT OSs like Mbed~\cite{mbed}, RIOT OS~\cite{riot_os}, Contiki-NG~\cite{contiki_ng}, and many others, with a strong focus on IoT device security~\cite{pop00010}.
+LwM2M has been integrated on most IoT OSs like FreeRTOS~\cite{freertos}, Mbed~\cite{mbed}, RIOT OS~\cite{riot_os}, Contiki-NG~\cite{contiki_ng}, and many others, with a strong focus on IoT device security~\cite{pop00010}.
+
+LwM2M has also proven to be a preferred solution for managing cellular IoT (CIoT) devices~\cite{10333286} due to its low overhead and comprehensive management capabilities, particularly in environments where network efficiency, scalability, and efficient firmware updates are critical~\cite{8725488}.
 
@@
 
-LwM2M, though robust in its core functionalities, has advanced significantly through various extensions and integrations.
-This section explores recent research developments that may inspire its future evolution.
-LwM2M has become a central focus in IoT research due to its improved security, in particular for firmware updates ~\cite{8725488} and integration with other technologies.
+LwM2M, though robust in its core functionalities, has advanced through various extensions and integrations.
+This section explores a subset of recent research developments that may inspire the future evolution of the standard.
 
 Several studies have concentrated on enhancing IoT security through the use of LwM2M.
-For instance, Muhammad et al.~\cite{pop00001} investigated the use of ARIA cryptography within Hardware Secure Modules for both LwM2M and MQTT protocols, aiming to bolster the security of next-generation IoT systems.
-Similarly, Lanzieri et al.~\cite{pop00011} proposed extensions to the LwM2M core specification to facilitate secure and authorized client-to-client communication, thereby addressing a significant limitation of the current standard.
+For instance, Muhammad et al.~\cite{pop00001} investigated the use of ARIA cryptography within Hardware Secure Modules for both LwM2M and MQTT protocols.
+Similarly, Lanzieri et al.~\cite{pop00011} proposed extensions to the LwM2M core specification to facilitate secure and authorized client-to-client communication, which is a limitation of the current standard.
 
 Others have focused on interoperability between SDOs, Kim et al.~\cite{pop00007} have designed and implemented a blockchain-based system that enables interworking between oneM2M~\cite{oneM2M-spec} and LwM2M IoT systems.
-This approach not only improves interoperability between different IoT frameworks but also enhances security by leveraging blockchain's immutable and distributed ledger properties.
+Their approach uses blockchain's immutable and distributed ledger properties for interoperability and security.
 
 In the context of Industrial IoT (IIoT), Yaker et al.~\cite{pop00004} introduced a novel edge Security Information and Event Management (SIEM) system for managing IoT flows within 5G private networks.
-Their approach incorporates LwM2M data events, demonstrating how edge computing can effectively manage and secure IoT data in a 5G environment.
+Their approach incorporates LwM2M data events to manage and secure IoT data in a 5G environment.
 Similarly, Myoung et al.~\cite{pop00012} addressed the integration of LwM2M with smart metering technologies, proposing a data interworking model between the Device Language Message Specification (DLMS)~\cite{dlms-spec} used for managing smart meters and the LwM2M protocol.
-This highlights the potential for harmonizing different IoT protocols within metering infrastructure systems.
 
@@
 Karaagac et al.~\cite{pop00013} explored the interoperability between LwM2M and the Open Platform Communications Unified Architecture (OPC UA), proposing a framework where OPC UA Servers can be virtualized as LwM2M Clients and vice versa.
-This approach aims to bridge the gap between IoT and industrial automation, facilitating seamless communication across heterogeneous systems.
-Similarly, Cavalcanti et al.~\cite{pop00014} reviewed various machine-to-machine communication protocols within the context of Industry 4.0, emphasizing the advantages of integrating OPC UA with LwM2M for industrial applications.
-
-A last example, Wang et al.~\cite{pop00015} proposed a CoAP-based OPC UA transmission scheme tailored for resource-constrained devices.
-Although their work does not directly involve LwM2M, it illustrates the broader trend of adapting industrial communication protocols to fit the needs of IoT environments and the reuse of IETF protocols in wider contexts that originally intended.
-This could also indicate the interest in enabling resource-efficient and interoperable solutions in the IoT landscape, where industrial and IoT domains converge.
+Similarly, Cavalcanti et al.~\cite{pop00014} reviewed various machine-to-machine communication protocols within the context of Industry 4.0.
+A third example is based only on CoAP, Wang et al.~\cite{pop00015}, which proposed a CoAP-based OPC UA transmission scheme tailored for resource-constrained devices.
+These three papers aim to bridge the gap between IoT and industrial automation.
 
@@
 
-In this paper, we have presented an overview of LwM2M and its components, the dependencies on IETF standards and some potential new building blocks for future additions.
+This paper provides an overview of the LwM2M protocol, detailing its core components, architecture, and integration with IETF standards.
+We explored its role in managing IoT devices, including its communication interfaces and data model, and discussed its adaptability through recent extensions.
+LwM2M continues to evolve as a key solution for IoT device management, offering scalability, security, and efficient operation, especially in constrained environments like Cellular IoT (CIoT).
+Future developments in interoperability and security enhancements will further strengthen its utility across diverse IoT applications.
 
@@
 
-We'd like to thank Ericsson for their support of this work.
+We would like to thank Ericsson for their support of this work. Special thanks to Jari Arkko for his valuable review and feedback, which improved the quality of this paper.
 

### 2024-10-28 paper.tex
cabo-review

@@
 
-%Jaime: We need to address gaps, such as operator-specific challenges or more detailed recommendations for future standardization efforts.
-
 The rapid evolution of network management protocols necessitates a reevaluation of existing technologies and their applicability to modern challenges.
@@
 The \hyperref[overview]{LwM2M Protocol Overview} details its architecture, focusing on communication between Clients, Servers, and Bootstrap Servers (see Figure~\ref{fig:overall_architecture}).
-Recent advancements, including integrations with blockchain and industrial protocols, are discussed in \hyperref[extensions]{LwM2M Extensions}, the \hyperref[conclusions]{Conclusions} summarize the main points.
+Recent advancements, including integrations with blockchain and industrial protocols, are discussed in \hyperref[extensions]{LwM2M Extensions}, the \hyperref[conclusions]{Conclusions} summarize the main points of the paper.
 
@@
 For secure communications, LwM2M often relies on the \textit{Datagram Transport Layer Security (DTLS)} as outlined in \textit{RFC 6347}~\cite{rfc6347}, ensuring encryption and integrity over the CoAP protocol.
-An additional layer of security is provided by \textit{Object Security for Constrained RESTful Environments (OSCORE)}~\cite{rfc8613}, which offers end-to-end encryption and integrity protection directly at the application layer, making it suitable for scenarios where DTLS is not applicable.
-Additionally, the \textit{Constrained RESTful Environments (CoRE) Resource Directory (RD)}~\cite{rfc9176} facilitates resource registration of IoT endpoints by maintaining information about resources on other servers, the lookup interface, although present is not intended for applications but for management purposes.
+An alternative layer of security is provided by \textit{Object Security for Constrained RESTful Environments (OSCORE)}~\cite{rfc8613}, which offers end-to-end encryption and integrity protection directly at the application layer, making it suitable for scenarios where DTLS is not applicable.
+Lastly the \textit{Constrained RESTful Environments (CoRE) Resource Directory (RD)}~\cite{rfc9176} facilitates resource registration of IoT endpoints.
+It does so by maintaining information about resources on other servers, the lookup interface, although present is not intended for applications but solely for management purposes.
 \sloppy
 
-Furthermore, LwM2M supports additional transport protocols, including the \textit{Hypertext Transfer Protocol (HTTP)}~\cite{http-spec} and \textit{Message Queuing Telemetry Transport (MQTT)}~\cite{mqtt-spec}, the latter developed by the OASIS foundation~\cite{oasis-foundation}.
-This broadens its applicability across various network environments and use cases.
+LwM2M also supports additional transport protocols, such as IETF's \textit{Hypertext Transfer Protocol (HTTP)}\cite{http-spec} and OASIS' \textit{Message Queuing Telemetry Transport (MQTT)}\cite{mqtt-spec}, expanding its applicability across diverse network environments and use cases.
 
@@
 The format is as follows:
-
 \[
@@
 \]
-
 In this structure, the instance or resource ID can be omitted if the request targets the entire object or a specific instance.
@@
 
-Several studies have concentrated on enhancing IoT security through the use of LwM2M.
+Firstly, LwM2M has developed a Northbound Web API to simplify interaction with LwM2M systems by abstracting the protocol's complexities~\cite{lwm2m-northbound-api}.
+This API enables external applications to perform key operations such as tracking device registrations, executing management commands (e.g., read, write, delete), and subscribing to observations for real-time monitoring of device resources.
+
+Outside OMA, several studies have concentrated on enhancing IoT security through the use of LwM2M.
 For instance, Muhammad et al.~\cite{pop00001} investigated the use of ARIA cryptography within Hardware Secure Modules for both LwM2M and MQTT protocols.
@@
 
-We would like to thank Ericsson for their support of this work. Special thanks to Jari Arkko for his valuable review and feedback, which improved the quality of this paper.
+We would like to thank Ericsson for their support of this work. Special thanks to Jari Arkko and Carsten Bormann for their valuable review and feedback, which improved the quality of this paper.
 

### 2024-11-12 paper.tex
incorporated sandor's feedback

@@
 
+The evolution of LwM2M has also prompted considerations for enhancing its data model capabilities.
+While the current IPSO data model, utilizing the format \textit{/<ObjectID>/<InstanceID>/<ResourceID>}, is well-suited for constrained environments, it may fall short in addressing the needs of complex machines or network nodes with large, deeply nested configuration structures.
+This limitation suggests the potential for developing alternative data models that can accommodate more intricate configurations, thereby extending the LwM2M protocol's applicability to a broader range of devices.
+
+The interoperability of LwM2M with hyperscalers is hindered by the lack of native CoAP support within these platforms.
+This limitation poses a challenge for vendors seeking to integrate LwM2M-based IoT solutions with HCS services.
+To overcome this obstacle, some vendors have devised methods to encapsulate CoAP messages within MQTT topics, although this solution is merely a workaround that may not fully address the efficiency needs of large-scale IoT deployments.
+
 \section{Conclusions} \label{conclusions}
@@
 
-We would like to thank Ericsson for their support of this work. Special thanks to Jari Arkko and Carsten Bormann for their valuable review and feedback, which improved the quality of this paper.
+We would like to thank Ericsson for their support of this work.
+Special thanks to Jari Arkko, Carsten Bormann and Sándor Katona for their valuable review and feedback, which improved the quality of this paper.
 
